New lower bounds for Schur and weak Schur numbers

### 2021-04-07 article/weakschur.tex

\documentclass{article}
\title{New theoretical lower bounds of weak Schur numbers using regular Schur numbers}
\author{Romain Ageron, Paul Casteras, Thibaut Pellerin, Yann Portella}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage[top=1cm]{geometry}
\usepackage{graphicx}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\newtheorem{definition}{Definition}[section]
\newtheorem{notation}[definition]{Notation}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]

\begin{document}
\maketitle
\section{Introduction, context and notations}
We start by defining sum-free and weakly sum-free subsets to introduce regular and weak Schur numbers.
\begin{definition}
A subset $A$ of $\mathbb{N}$ is said to be sum-free when :
\[ \forall (a,b) \in A^2 \text{, } a+b \notin A
\end{definition}
\begin{definition}
A subset $B$ of $\mathbb{N}$ is said to be weakly sum-free when :
\[ \forall (a,b) \in B^2 \text{, } a \neq b \Longrightarrow a+b \notin B
\end{definition}
Let us notice that a sum-free subset is also weakly sum-free, hence justifying the name of \textit{weakly} sum-free subsets. Given $m$ and $n$ two integers, we are interested in partitioning the set of integers from 1 to $m$ in $n$ weakly sum-free subsets.
\begin{notation}
We denote by $[\![1,n]\!]$ the set of integers $\{1, 2, ..., n\}$
\end{notation}
Schur proved in \hyperlink{label1}{\textbf{[1]}} that given a number of subsets $n$, there exists a value of $m$ such that there exists no partition of $[\![1,p]\!]$ in $n$ sum-free subsets for any $p \geqslant m$. This observation leads to the following definition.
\begin{definition}
Let $n \in \mathbb{N}^*$, there exists an integer that we note S(n) (\textit{respectively WS(n)}) such that $[\![1,S(n) + 1]\!]$ (respectively $[\![1,WS(n) + 1]\!]$) can not be partitioned in $n$ sum-free subsets (respectively in $n$ weakly sum-free subsets) while $[\![1,S(n)]\!]$ (respectively $[\![1,WS(n)]\!]$) can be. S(n) is called the n\textsuperscript{th} Schur number and WS(n) the n\textsuperscript{th} weak Schur number.
\end{definition}
The values of both Schur and weak Schur numbers are known only for small values of $n$, and our goal is to improve the lower bounds on weak Schur numbers for $n \geqslant 5$. At the time being, the best lower bounds for weak Schur numbers are established by Rowley in \hyperlink{label2}{\textbf{[2]}}. He greatly improved the former best values with a purely theoretical and constructive approach. More precisely, he managed to build interesting weakly sum-free partitions with sum-free partitions. With this, he found the following lower bounds :
\[ WS(n+1) \geqslant 4S(n) + 2
\[
WS(n+2) \geqslant 13S(n) + 8
Actually, Rowley's construction can be generalised to find a lower bound for $WS(n+m)$ for all $(n,m) \in (\mathbb{N}^*)^2$. This result and its demonstration are detailled in the next part.
\section{A new theoretical lower bound for WS(n+m)}

@@

Let $(n,p), (r,s) \in (\mathbb{N}^*)^2$. If there exists a partition of $r$ weakly sum-free subsets of $[\![1,n]\!]$ and a partition of $s$ sum-free subsets of $[\![1,p]\!]$ then there exists a partition of $r+s$ weakly sum-free subsets of $[\![1,p(n+\left \lfloor \frac{n}{2} \right \rfloor + 1)+n]\!]$

@@

$ \forall (n,m) \in (\mathbb{N}^*)^2 \text{, } WS(n+m) \geqslant S(m) \left (WS(n) + \left \lfloor \frac{WS(n)}{2} \right \rfloor +1 \right) + WS(n)$

@@

\textsc{Proof :} Let $(n,p), (r,s) \in (\mathbb{N}^*)^2$,  $N = p(n+\left \lfloor \frac{n}{2} \right \rfloor + 1)+n$, $\alpha = \left \lfloor \frac{n}{2} \right \rfloor > 0$ and $\beta = n + \alpha + 1$. We denote by $f$ the projection of the equivalence relation induced by the partition of $[\![1,n]\!]$ and $g$ the one induced by the partition of $[\![1,p]\!]$. Each equivalence class is represented by a single integer, therefore :

@@

\section{References}
\hypertarget{label1}{\textbf{[1]}} Schur, I.. "Über Kongruenz x ... (mod. p.).." Jahresbericht der Deutschen Mathematiker-Vereinigung 25 (1917): 114-116. \url{http://eudml.org/doc/145475} \\
\hypertarget{label2}{\textbf{[2]}} F. Rowley, New Lower Bounds for Weak Schur Partitions, arXiv 2011.11292, \url{https://arxiv.org/pdf/2011.11292.pdf}
\end{document}

>>>>

\section{Weak Schur numbers}

In this section, we generalize Rowley's constructions in [2]. We then introduce, by analogy with the third section, the integer \(WE_+(n)\) 
to build suitable templates.

\subsection{Lower bound for Weak Schur numbers using Schur and Weak Schur numbers}

@@

	Let $(n,p), (r,s) \in (\mathbb{N}^*)^2$. If there exists a partition of $r$ weakly sum-free subsets of $[\![1,n]\!]$ and a partition of $s$ sum-free 
	subsets of $[\![1,p]\!]$ then there exists a partition of $r+s$ weakly sum-free subsets of $[\![1,p(n+\left \lceil \frac{n}{2} \right \rceil + 1)+n]\!]$

@@

	$ \forall (n,m) \in (\mathbb{N}^*)^2 \text{, } WS(n+m) \geqslant S(m) \left (WS(n) + \left \lceil \frac{WS(n)}{2} \right \rceil +1 \right) + WS(n)$

@@

\textsc{Proof :} Let $(n,p), (r,s) \in (\mathbb{N}^*)^2$,  $N = p(n+\left \lceil \frac{n}{2} \right \rceil + 1)+n$, $\alpha = \left \lceil \frac{n}{2} \right \rceil > 0$ and $\beta = n + \alpha + 1$. We denote by $f$ the projection of the equivalence relation induced by the partition of $[\![1,n]\!]$ and $g$ the one induced by the partition of $[\![1,p]\!]$. Each equivalence class is represented by a single integer, therefore :

@@
